---
title: "OLSengine: Assisted Simplicity Tutorial"
author: "msoto-perez"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{OLSengine Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# 1. Introduction

The philosophy of **Assisted Simplicity** aims to bridge the gap between complex computation and methodological rigor. **OLSengine** provides six estimation engines in a single, unified interface: OLS regression, ANOVA/t-tests, logistic regression, panel data, instrumental variables (2SLS), and difference-in-differences.

All methods are implemented in pure base R with integrated **Methodological Customs** that audit assumptions and guide researchers toward robust alternatives.

---

# 2. OLS Regression: Handling Heteroskedasticity

In social sciences, variance often increases with the scale of the predictor. Let's see how `OLSengine` detects and handles this.

```{r, message=FALSE}
library(OLSengine)

# Simulate data with non-constant variance
set.seed(123)
n <- 200
x <- rnorm(n, 50, 10)
y <- 10 + 0.5 * x + rnorm(n, 0, x * 0.2) # Heteroskedasticity

df <- data.frame(y, x)

# Run the engine
model <- paper_engine(y ~ x, data = df, model = "ols")
model$messages
```

### Correcting with Robust Standard Errors

Following the Aduana's advice, we apply HC3 robust standard errors:

```{r}
model_robust <- paper_engine(y ~ x, data = df, model = "ols", robust = TRUE)
model_robust$tables$Table2_OLS_Estimation
plot_engine(model_robust)
```

---

# 3. ANOVA/t-tests: Experimental Comparisons

In experimental research, we often compare groups. `OLSengine` automatically checks normality and variance homogeneity.

```{r}
# Simulating 3 groups with non-normal distribution
set.seed(789)
group_data <- data.frame(
  score = c(rgamma(30, 2, 0.5), rgamma(30, 5, 0.5), rgamma(30, 3, 0.5)),
  group = rep(c("Control", "Treatment A", "Treatment B"), each = 30)
)

# Auto-pilot switches to non-parametric if normality fails
model_anova <- paper_engine(score ~ group, data = group_data, 
                            model = "anova", non_parametric = "auto")

model_anova$tables$Table4_Mean_Differences
plot_engine(model_anova)
```

---

# 4. Logistic Regression: Binary Outcomes

When the outcome is binary (e.g., Success/Failure), the engine calculates Odds Ratios, McFadden's and Nagelkerke's Pseudo R², and classification accuracy.

```{r}
# Simulating binary data
set.seed(101)
n_logit <- 100
age <- rnorm(n_logit, 40, 10)
passed <- rbinom(n_logit, 1, plogis(-5 + 0.12 * age))

logit_df <- data.frame(passed, age)

# Run Logit engine
model_logit <- paper_engine(passed ~ age, data = logit_df, model = "logit")

model_logit$tables$Table2_Logit_Estimation
model_logit$messages
plot_engine(model_logit)
```

---

# 5. Panel Data: Fixed and Random Effects

Panel data tracks the same entities over time. The Hausman test automatically selects between Fixed Effects (FE) and Random Effects (RE).

```{r}
# Simulate panel data: 50 workers observed over 3 years
set.seed(456)
n_entities <- 50
n_time <- 3

panel_df <- data.frame(
  worker_id = rep(1:n_entities, each = n_time),
  year = rep(2018:2020, times = n_entities),
  wage = rnorm(n_entities * n_time, 50, 10) + 
         rep(rnorm(n_entities, 0, 5), each = n_time), # Entity fixed effect
  experience = rep(5:7, times = n_entities) + rnorm(n_entities * n_time, 0, 1)
)

# Hausman test decides between FE and RE
model_panel <- paper_engine(wage ~ experience, 
                            data = panel_df,
                            model = "panel",
                            entity_id = "worker_id",
                            time_id = "year",
                            method = "auto")

model_panel$tables[[1]]
model_panel$messages
plot_engine(model_panel)
```

---

# 6. Instrumental Variables: Addressing Endogeneity

When a predictor is correlated with the error term (endogeneity), Instrumental Variables (IV) provide consistent estimates using Two-Stage Least Squares (2SLS).

```{r}
# Simulate IV data with endogeneity
set.seed(789)
n_iv <- 200

z <- rnorm(n_iv, 10, 3)  # Instrument
u <- rnorm(n_iv, 0, 2)   # Unobserved confounder

x <- 2 + 0.8 * z + u + rnorm(n_iv, sd = 1)  # Endogenous predictor
y <- 5 + 1.5 * x + u + rnorm(n_iv, sd = 3)  # Outcome

iv_df <- data.frame(y, x, z)

# 2SLS estimation
model_iv <- paper_engine(y ~ x, 
                         data = iv_df,
                         model = "iv",
                         instruments = ~ z)

model_iv$tables$Table2_IV_2SLS
model_iv$messages  # Reports first-stage F-stat
plot_engine(model_iv)
```

---

# 7. Difference-in-Differences: Causal Policy Evaluation

DiD estimates treatment effects by comparing changes over time between treated and control groups.

```{r}
# Simulate policy intervention data
set.seed(321)
n_per_group <- 100

did_df <- data.frame(
  outcome = c(rnorm(n_per_group, 50, 10), rnorm(n_per_group, 53, 10),  # Control: pre/post
              rnorm(n_per_group, 50, 10), rnorm(n_per_group, 58, 10)), # Treated: pre/post
  group = rep(c("Control", "Treated"), each = n_per_group * 2),
  period = rep(c("Pre", "Post"), times = n_per_group * 2)
)

# DiD estimation
model_did <- paper_engine(outcome ~ 1, 
                          data = did_df,
                          model = "did",
                          treatment_var = "group",
                          time_var = "period",
                          treatment_level = "Treated",
                          post_level = "Post")

model_did$tables$Table2_DiD_Estimation
model_did$tables$Group_Means
model_did$messages  # Reports parallel trends test
plot_engine(model_did)
```

---

# 8. Real Data Example: Academic Salaries

The package includes a real dataset of 397 U.S. college professors.

```{r}
data(academic_salaries)

# Explore salary determinants
salary_model <- paper_engine(salary ~ rank + discipline + years_since_phd + sex,
                             data = academic_salaries,
                             model = "ols",
                             robust = "auto")

salary_model$tables$Table2_OLS_Estimation
salary_model$messages
```

---

# Conclusion

`OLSengine` simplifies the transition from raw data to paper-ready results across six estimation methods. Every step is backed by rigorous methodological audits, ensuring researchers can focus on interpretation rather than implementation.

**Key Features:**
- Zero external dependencies (pure base R)
- Integrated diagnostics with actionable warnings
- APA-formatted tables and publication-ready plots
- Unified interface across OLS, ANOVA, Logit, Panel, IV, and DiD
